---
title: "Class 7: Machine learning 1"
author: "Jie"
format: pdf
---

# K-means Clustering

```{r}
tmp <- c(rnorm(30,-3),rnorm(30,3))
x <- cbind(x=tmp,y=rev(tmp))
plot(x)
```

```{r}
km <- kmeans(x,centers=4,nstart=20)
km
```

```{r}
km$size
km$cluster
km$centers
```

```{r}
plot(x,col=km$cluster)
points(km$centers,col=km$clusters,pch=15,cex=2)

#rep(x,y) repeat membership x for y times
```

```{r}
hc <- hclust(dist(x))
hc
plot(hc)
```

```{r}
cutree(hc,h=8)
```

```{r}
grsp <- cutree(hc,k=2)
grsp
```

```{r}
plot(x,col=grsp)
```

#Principal Component Analysis (PCA)

Q2. Which approach to solving the ‘row-names problem’ mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url,row.names=1)
```

```{r}
head(x)
```

```{r}
#rownames(x) <- x[,1]
#x <- x[,-1]
#x
```

Q1. How many rows and columns are in your new data frame named x? What R functions could you use to answer this questions?

```{r}
dim(x)
```

Q3: Changing what optional argument in the above barplot() function results in the following plot?

```{r}
barplot(as.matrix(x),col=rainbow(nrow(x)))
```


Q5: Generating all pairwise plots may help somewhat. Can you make sense of the following code and resulting figure? What does it mean if a given point lies on the diagonal for a given plot?

```{r}
pairs(x, col=rainbow(10), pch=16)
```

Q6. What is the main differences between N. Ireland and the other countries of the UK in terms of this data-set?

```{r}
#prcomp expects the transpose of the data input
pca <- prcomp(t(x))
summary(pca)
```

Q7. Complete the code below to generate a plot of PC1 vs PC2. The second line adds text labels over the data points.

```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
```

Q8. Customize your plot so that the colors of the country names match the colors in our UK and Ireland map and table at start of this document.

```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x),col=c("orange","red", "blue","darkgreen"))
```

```{r}
par(mar=c(10, 3, 0.35, 0))
barplot(pca$rotation[,1], las=2)
```
Q9: Generate a similar ‘loadings plot’ for PC2. What two food groups feature prominantely and what does PC2 maninly tell us about?

```{r}
par(mar=c(10, 3, 0.35, 0))
barplot(pca$rotation[,2], las=2)
```

```{r}
biplot(pca)
```



