---
title: "Hawollen candy project!"
author: "Jie"
format: pdf
---

# 1. Importing candy data

```{r}
candy <- read.csv("candy-data.csv", row.names=1)
head(candy)
```

Q1. How many different candy types are in this dataset?
85

```{r}
nrow(candy)
```

```{r}
rownames(candy)
```
Q2. How many fruity candy types are in the dataset?
38

```{r}
sum(candy$fruity)
```

# 2. What is your favorate candy?

Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}

candy["Haribo Gold Bears", ]$winpercent

```

Q4. What is the winpercent value for “Kit Kat”?

```{r}

candy["Kit Kat", ]$winpercent

```
Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?


```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```


```{r}
library("skimr")
skim(candy)
```
Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?
Winpercent looks diffent as its mean is around 50 while others are less than 1

Q7. What do you think a zero and one represent for the candy$chocolate column?
0 is the candy does not contain chocolate and 1 is the candy does contain chocolate

Q8. Plot a histogram of winpercent values

```{r}

hist(candy$winpercent)

```

```{r}
library("ggplot2")

ggplot(candy) + aes(winpercent) + geom_histogram(bins=10,col="yellow")
```

Q9. Is the distribution of winpercent values symmetrical?
No

Q10. Is the center of the distribution above or below 50%?
Below 50%

Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
chocolate.inds <- as.logical(candy$chocolate)
chocolate.win <- candy[chocolate.inds,]$winpercent
mean(chocolate.win)
```

```{r}
fruity.inds <- as.logical(candy$fruity)
fruity.win <- candy[fruity.inds,]$winpercent
mean(fruity.win)
```
Chocolate is lower ranked than fruit candy 

Q12. Is this difference statistically significant?

```{r}
t.test(chocolate.win,fruity.win)
```
It is significant


# 3. Overall Candy Rankings

Q13. What are the five least liked candy types in this set?

```{r}
#x[ order(x) ]
head(candy[order(candy$winpercent),], n=5)
```

Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
tail(candy[order(candy$winpercent),], n=5)
```

Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_bar(stat='identity')
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_bar(stat='identity')
```

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

- Q17. What is the worst ranked chocolate candy?
Sixlets

- Q18. What is the best ranked fruity candy?
Starburst

# 4. Taking a look at pricepercent

```{r}
library(ggrepel)

ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?
ReeseOs Miniatures


Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?
Nik L Nip

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

# 5. Exploring the correlation structure

```{r}
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```

Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?
Chocolate and fruity 

Q23. Similarly, what two variables are most positively correlated?
Chocolate and bar

# 6. Principal Component Analysis

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=2.2, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```


```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?
Fruity, hard and pluribus