---
title: "HW6"
output: pdf_document
date: "2022-10-16"
---

```{r}
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

Q1. What type of object is returned from the read.pdb() function?

```{r}
class(s1)
```

**PDB structure object**

Q2. What does the trim.pdb() function do?

**It reduces pdb to a smaller pdb object that only contains a subset of chain and type**

Q3. What input parameter would turn off the marginal black and grey rectangles in the plots and what do they represent in this case?

**sse parameter would turn off the rectangles and they represent secondary structure elements(black: alpha helices; light grey: beta sheets)**

Q4. What would be a better plot to compare across the different proteins?

**Plots that can show sequence and structural conservation, for example structural superposition diagram or alignment graphs**

Q5. Which proteins are more similar to each other in their B-factor trends. How could you quantify this?

```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

**s2 and s3 are more similar, height is used to quantify them also the joint between s2 and s3 show that they are more closely related**

## Write a function

Q6. How would you generalize the original code above to work with any set of input protein structures?

```{r}

pdbanalysis <- function(x){
    s <- read.pdb(x) 
    s.chainA <- trim.pdb(s, chain="A", elety="CA")
    s.b <- s.chainA$atom$b
    plotb3(s.b, sse=s.chainA, typ="l", ylab="Bfactor")
    return(s.b)
    
  }
```

1.  Input of the function is the pdb structure object(pdb file)
2.  The function is to read the pdb file, reduce it to a specified and desired chain to analyze a given kinase
3.  The output is a plot of residues vs B-factor with noted secondary structure elements

```{r}
#Example output
pdbanalysis ("4AKE")
pdbanalysis ("1AKE")
pdbanalysis ("1E4Y")
```
